Towards a Pandemic-Scale Ancestral Recombination Graph of SARS-CoV-2

### 2023-05-12 paper.tex
Yan's updates

\subsection{Backbone phylogeny}
\label{sec:backbone_phylogeny}
For validation purposes, we compare the genetic
relationships in our ARGs with phylogenetic relationships
inferred by the NextStrain project \citep{Hadfield2018-ef}.
To enable this comparison, we ``simplified''~\citep{Kelleher2018-xc}
the sc2ts ARGs to a ``backbone'' containing only samples present
in a specific NextStrain tree. For this purpose, we picked an
``all-data''  NextStrain tree restricted to the time period covered
by the appropriate ARG. This results in a relatively small
set of shared samples, almost entirely restricted to non-recombinant lineages.
These simplified backbone ARGs exhibit very
little recombination (6 and 5 trees across the genome for the Wide and Long
ARGs, respectively).
% JK: This leaves a big question hanging - why do we ahave recombination
% in here at all? It's probably just artefactual, right?

>>>>

\subsection{Backbone phylogeny}
\label{sec:backbone_phylogeny}
%% The ARG is mostly tree-like
Compared to organisms like humans which recombine every generation,
SARS-CoV2 shows little recombination, with recombination nodes accounting
for <0.3\% of our inferred ancestry (Table~\ref{tab:args}). As a result,
relationships between strains can often be adequately represented by a
single tree, particularly early in the pandemic, or when looking at
a subset of sampled genomes e.g. restricted by time or limited to close
relatives.

For validation purposes, we compare the genetic
relationships in our ARGs with phylogenetic relationships
inferred by the NextStrain project \citep{Hadfield2018-ef}.
To enable this comparison, we ``simplified''~\citep{Kelleher2018-xc}
the sc2ts ARGs to a ``backbone'' containing only samples present
in a specific NextStrain tree. For this purpose, we picked an
``all-data''  NextStrain tree restricted to the time period covered
by the appropriate ARG. This results in a relatively small
set of shared samples, almost entirely restricted to non-recombinant lineages.
These simplified backbone ARGs exhibit very
little recombination (6 and 5 trees across the genome for the Wide and Long
ARGs, respectively).
% JK: This leaves a big question hanging - why do we ahave recombination
% in here at all? It's probably just artefactual, right?
